Accelerating the Computation of Geometric Constraints in Optimization

### 2019-12-11 report/paper.tex

When one object encloses another, the minimum distance $d_\text{min}$ between them is greater than zero by some margin.
Therefore, a first approach to a geometric constraint might involve computing $d_\text{min,AB} \geq 0 + \text{tol}$.
We can compute $d_\text{min}$ between two triangulated surfaces by computing the minimum distance between each individual pair of triangles.

@@

The minimum distance between a pair of triangles can be found by a total of fifteen primitive tests between the vertices: six point-triangle tests, and nine line-line tests. % TODO cite ericson

\begin{equation}
  d_{\text{min},ij} = \text{min}(d_{\text{min},k}) \text{ for primitive test }k\text{1 to 15}
\end{equation}

The distances are then aggregated into a single scalar and constrained for feasibility using the alternate form of the Kresselmeier-Steinhauser (KS) function ~\ref{ks_geom}.
\begin{equation}
  \label{ks_geom}
  KS_\text{geom}(\textbf{x}) = \frac{1}{\rho} \textrm{ln} \Bigg[\sum_{j=1}^{J} e^{\rho\big(d_\text{min}\:(\textbf{x})-d_j\:(\textbf{x})\big)}\Bigg] - d_\text{min}\:(\textbf{x}) \leq 0 ,
\end{equation}

where $\textbf{x}$ is a vector of design variables,
$d_j(\textbf{x})$ is the $j\textrm{th}$ pairwise distance between $A$ and $B$,
$d_{min}(\textbf{x})$ is the minimum distance between $A$ and $B$, and
$\rho$ is a user-controllable constant.

>>>>

When one object encloses another, the minimum distance $d_\text{min}$ between them is greater than zero by some margin.
Therefore, a first approach to a geometric constraint might involve computing $d_\text{min,AB} \geq 0 + \text{tol}$.
We can compute $d_\text{min}$ between two triangulated surfaces by computing the minimum distance between each individual pair of triangles.
The minimum distance between a pair of triangles can be found by a total of fifteen primitive tests between the vertices: six point-triangle tests, and nine line-line tests. % TODO cite ericson

@@

During optimization, the pair of facets comprising the minimum distance can change rapidly, leading to discontinuities in the gradients and poor optimizer performance.
Notionally, we want to gather information not only from the exactly closest triangles, but also from the triangles that are nearly closest.
We can achieve this by using a strategy called constraint aggregation.
I used a function called the Kreisselmeier-Steinhauser (KS) function which essentially provides a conservative estimate of the extremum of a vector.
The computation is as follows:

\begin{equation}
  \label{ks_geom}
  KS_\text{geom}(\textbf{x}) = \frac{1}{\rho} \textrm{ln} \Bigg[\sum_{i,j=1}^{m,n} e^{\rho\big(d_\text{min}\:(\textbf{x})-d_{ij}\:(\textbf{x})\big)}\Bigg] - d_\text{min}\:(\textbf{x}) \leq 0 ,
\end{equation}

where $\textbf{x}$ is a vector of design variables,
$d_{ij}(\textbf{x})$ is the pairwise distance between the $i$th facet of $A$ and $j$th facet of $B$,
$d_{min}(\textbf{x})$ is the minimum distance between $A$ and $B$ at the current design point, and
$\rho$ is a user-controllable constant.
